---
title: "Manner of Exercise Predictions"
author: "Miguel Alfonso Bustamante Sanchez"
date: "Jul 2, 2017"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

The analysis shows that Manual Transmission has better mpg than Automatic Transmission, this holds with the other predictors weight (wt) and quarter mile time (qsec).
The mean MPG difference is 7.25 with an averga of 17.14 for automatic transmission and 24.39 for manual transmission.

We can conclude that manual transmission is better for mpg.

## Data

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

```{r}
library(caret)
library(knitr)
library(rpart.plot)

training <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings = c("NA", "", "#DIV/0"))

testing <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings = c("NA", "", "#DIV/0"))

dim(training)
dim(testing)


inTrain <- createDataPartition(training$classe, p=0.6, list = FALSE)
training2 <- training[inTrain,]
testing2 <- training[-inTrain,]

dim(training2)
dim(testing2)

training2 <- training2[lapply(training2, function(x) sum(is.na(x))/length(x))<0.6]

colSToClean <- colnames(training2)

testing2 <- testing2[colSToClean]
```


```{r}
set.seed(12345)
modelDT <- train(classe ~ .,data = training2, method="rpart", tuneLength=10)
plot(modelDT$finalModel)


predictDT <- predict(modelDT, testing2, type = "raw")
cDT <- confusionMatrix(predictDT, testing2$classe)
cDT
```

```{r}
set.seed(12345)
modelRF <- train(classe ~ .,data = training2, method="rf", tuneLength = 10)
plot(modelRF$finalModel)


predictRF <- predict(modelRF, testing2, type = "raw")
cRF <- confusionMatrix(predictRF, testing2$classe)
cRF
```
```{r}
set.seed(12345)
modelGBM <- train(classe ~ .,data = training2, method="gbm", tuneLength = 10)
plot(modelGBM$finalModel)


predictGBM <- predict(modelGBM, testing2, type = "raw")
mGBM <- confusionMatrix(predictGBM, testing2$classe)
mGBM
```
